### 2016-08-11 robust-software.tex

As a relatively young field, bioinformatics is full of newly developed software.
\fixme{MT}{examples here}
Efforts such as
\withurl{the ELIXIR tools and data registry}{http://dx.doi.org/10.1093/nar/gkv1116}
and the
\withurl{Bioinformatics Links Directory}{https://bioinformatics.ca/links\_directory/}
\cite{brazas2012}
have made efforts towards cataloguing it: as of May 2016, the former
has 2500 entries and the latter 1700, and those numbers are constantly
growing as both trainees and experienced practitioners produce new
software to support their research.

Typically, this software is initially developed by one person, and may
produce excellent results in their hands. But what happens after that
student leaves that lab and someone else wants to use it? Everyone with
a few years of experience feels a tremor of fear when told, ``Use
{\textless}graduated student{\textgreater}'s code to analyze your
data''. Often, that software will be undocumented and work in unexpected
ways (if it works at all without substantial modification). Equally
often, the potential new user winds up shaking their fist and cursing
the author's name. She then has two choices: hack the existing code to
make it work for her, or start over.

@@

\begin{itemize}
\item
  Is kept under version control.
\item
  Can be installed on systems outside the original institution
\item
  Works for users other than the original author
\item
  Has well-defined input and output formats
\item
  Has documentation that describes what its dependencies are, how to
  install it, and what the options are.
\item
  Comes with enough tests to show that it actually runs.
\end{itemize}

These are all necessary steps toward creating a reusable library that
can be shared through a site like CPAN or CRAN, and apply to both
closed-source and open-source software. They do not depend on specific
languages, libraries, packages, documentation styles, or operating
systems. Whether the aim is as simple as sharing the code with
collaborators or as complex as using the software in a production
analysis environment, increasing the robustness of your software
decreases headaches all around.

Note: we do not recommend that these rules be applied to \emph{every}
coding effort. The vast majority of code produced in the marathon of a
graduate thesis is ``throw-away'' code that is used once to answer a
specific question related to a specific dataset. However, once that
little script is dragged out three or four times for slightly different
purposes, it may be time to apply ten simple rules for robust software.
As the saying goes, not everything worth doing is worth doing right
away.

@@

The README is the first stop for any potential users interested in your
software. At a minimum, it needs to provide or point to everything a new
user needs to get started, where they can turn to for help, and which
licenses apply to the software package. Exhaustive details regarding
parameters and usage are not usually necessary in a README if they are
present in usage (\#2), although a working example using test data (per
\#9) is always appreciated.

\textbf{Explain what the software does:} At the beginning of the README,
explain what the software does in one or two sentences. The description
does not need to be long or detailed. There's nothing more frustrating
than spending the time to download and install some software only to
find out that it doesn't do what you thought it did.

\begin{verbatim}
Debarcer (De-Barcoding and Error Correction) is a package for
working with next-gen sequencing data that contains molecular
barcodes.
\end{verbatim}

\textbf{List required dependencies:} Often, software depends on very
specific versions of libraries, modules, or operating systems. This is
entirely reasonable as long as it is properly documented. Often,
multiple libraries exist with the same or very similar names, so
either provide the commands necessary to download the dependencies or
link to the software homepage. Include the version number for each
dependency.  Especially if you use older versions, include links where
the packages can be downloaded. Package managers like apt, pip and
homebrew stop offering older packages after a few years.
\fixme{MT}{Is this true of pip and Homebrew?}

If your dependency is to an internal package that is not available on
the internet, you have several options depending on the sensitivity of
the code in question. If it is plain text, you can add it directly to
you repository with appropriate attribution. If the dependency is a
binary, we recommend using a
\withurl{binary repository manager}{https://en.wikipedia.org/wiki/Binary\_repository\_manager}
such as Artifactory or ProGet. These managers keep versioned copies of
software at constant URLs so they can be downloaded as long as the
manager continues to run. As a last resort, you can place it at an
internal location on shared disk, remove all write permissions, and
link to it from your README, although this method is heavily
discouraged because of the potential for the directory to go missing
because of factors outside the developer's control.

\textbf{Installation instructions:} If the software needs to be compiled
or installed, list those instructions in the readme. New users may not
be familiar with your build system, even if it is \texttt{make}. Also
mention here if you recommend they use a pre-compiled binary instead
through a system such as pip or apt.

\textbf{Input and output files}: All possible input and output files
should be listed in this section. Do the files conform to a particular
industry standard, an extension of an existing format, or is it your own
format? If using a standard format, link to the specification and
version. If you extend the standard or have your own format, define it
here explicitly, listing all the required fields and acceptable values.
(You get bonus points if you include a script to convert between
standard format and your file format). If there is no rigorous format
(such as with log files), show an example file, or the first few lines,
and explain what the sections mean.

Input files and their formats are included in most documentation.
However, the definitions of the output files are often missing. In
addition to the expected output, software will often produce
intermediate files, auxilliary files, and log files. We believe
\emph{all} output files should be listed in the README. Log and
auxilliary files are often full of valuable information that can be
mined for the user's specific purpose. Even if the files are considered
self-explanatory. Sometimes your users will need to answer a question of
the format, ``Does X tell you the percentage of reads trimmed to remove
adapter sequences?'' and you can check the documentation and confidently
say ``yes, it is in the log file''.

\textbf{Attributions and licensing:} Attributions are how you credit
your main contributors; licenses are how you want others to use and
credit your software. Both are important in your README. Leave no
question in anyone's mind about whether your software can be used
commercially, how much modification is permitted, and how other software
needs to attribute to you. If your software is not open source, include
a statement here. Attributions can also contain a list of `expert' users
that can be contacted if new users have problems with the software. (for
better or for worse)

\begin{itemize}
\item
  This is readable \emph{before} the software is installed (or even
  downloaded).
\item
  Should also include (or better yet, point to) the license for the
  software, so that (potential) users will know what they're allowed to
  do.
\item
  \fixme{GW}{example from real software:\\ https://github.com/dib-lab/khmer/blob/master/README.rst}
\end{itemize}

@@

Users who run your software after installation may not have access your
well-crafted README (or may not have bothered to read it). Usage
information provides their first line of help.

Ideally, usage is a terse, informative command-line help message that
guides the user in the correct use of your software. Terseness is
important: usage that extends for multiple screens, especially when
printed to standard error instead of standard output (where it can
easily be paged), is a nuisance, and is as unlikely to be read as the
README file.

Usage should provide all of the information necessary to run the
software. It is usually invoked either by running the software without
any arguments; running the software with incorrect arguments; or by
explicitly choosing a help or usage option.

@@

\textbf{The syntax for running the program}: This defines the relative
location of optional and required flags and arguments for execution, and
includes the name of the program. Arguments in {[}square brackets{]}
tend to be optional. Multiple periods (e.g. ``{[}OPTION{]}\ldots{}'')
indicate that more than one can be provided.

\textbf{A text description of its purpose}: Similar to the README, the
description reminds users of the software's primary function.

\textbf{Most commonly used flags, a description of each flag, and the
default value}: Not all flags need to appear in the usage, but the most
commonly used ones should be listed here. Users will rely on this for
quick reference when working with your software.

@@

\textbf{Printed to standard output} : So that it can be piped into
\texttt{less}, searched with \texttt{grep}, or compared to the previous
version's help with \texttt{diff}.

\textbf{Exit with an appropriate exit code}: When usage is invoked by
providing incorrect parameters, the program should exit with a non-zero
code to indicate an error. However, when help is explicitly requested,
the software should not exit with an error, because requesting help is
sometimes used to verify that a dependency is available.

@@

Root (also known as ``superuser'' or ``admin'') is a special account on
a computer that has (among other things) the power to modify or delete
system files and user accounts. Conversely, files and directories owned
by root usually \emph{cannot} be modifed by normal users.

@@

You know what they say about assumptions*. As soon as you make one
assumption about how a software should work, a use case will come along
that will require you to change the assumption.

Every useful parameter should be configurable on the command line.
Useful parameters are those that a user will need to modify to suit
their computer, dataset or application. Providing parameters on the
command line increases the flexibility and usability of the program. You
may have determined early on that 0.58 is an optimal seed for your
original dataset, but that doesn't mean that is the best seed for every
case. Being able to change parameters on the fly to determine if and how
they change the results is important as your software gains more users,
facilitating exploratory analysis and parameter sweeping. Keep in mind
that if a parameter can be adjusted, users will want to be able to turn
off the feature entirely to have a baseline comparison.

The list of useful parameters is software-specific and so cannot be
provided here, but here is a short list of common useful parameters.

@@

@@

You can set reasonable default values to reduce the length of the
execution command as long as any parameters given on the command line
override those values. Configuration files should be used in preference
to hardcoding the defaults directly. Only values that are unlikely to
change between runs belong in the config file, such as dependencies,
servers, version numbers, network drives, and any other defaults for
your lab or institutions. Configuration files can be in a standard
location, e.g. \texttt{.packagerc} in the user's home directory or
provided on the command line as an additional argument.

We caution against overusing configuration files. When a user needs to
locate, open, change and save a file in order to change a parameter, the
import of the change seems larger and discourages experimentation. Since
all parameters should be echoed in the results, config files must be
cleaned up after execution completes or they occupy valuable disk space.
Use a default configuration file and specify all other parameters on the
command line.

* They make fools out of you and me.

@@

Hard-coding filenames in a program also makes the software harder to run
in other environments. If your package is installed on a cluster, for
example, the user's data will almost certainly \emph{not} be in the same
directory as the software, and the folder
\texttt{C:\textbackslash{}users\textbackslash{}yourname\textbackslash{}}
will probably not even exist.

@@

In the spirit of code reuse and interoperability, developers often want
to use software written by others. The tool could be standard for the
task at hand or do exactly what is necessary. With a few lines, a call
is made out to the other program, the results are incorporated into the
primary script. Using popular projects reduces the amount of code that
needs to be maintained and adds the strength of vetted software to the
final program.

@@

The way that the second program is invoked can throw up errors. For
example, the program may not be in the user's path, or it may be an
older or newer version and produce results differen than expected.
Windows users will be frustrated if you invoke bash or sh explictly or
use shell-specific conventions like `*' expansion. It is unlikely that
all of your users will all be on the operating system and version as
you. Even Linux-standard functions available vary slightly between
installs. For example: GNU \texttt{sort} is available on almost every
*nix distribution, but sorts differently depending on locale.

We fully support the reuse of software between projects and have some
suggestions to reduce the aforementioned pains. First, ensure the
appropriate software and version is available. Either allow the user to
configure the exact path to the package, distribute the program with the
dependent software, or download it during installation using a
dependency management system. Regardless, check whether the program is
executable and what version is running. Be sure to remind users in the
documentation what the compatible versions are.

Second, to ensure support on as many different operating systems as
possible, use native functions for starting other processes, such as
Java's Runtime.exec call, Python's subprocess module, and Perl's system
command, and be sure to capture and report the standard error output of
the subprocess to facilitate debugging.

Finally, make sure that you really need the call-out. If you are
executing GNU sort instead of figuring out how to sort lists in Python,
it may not be worth the tears of reuse.

@@

Many bioinformatics applications rely on randomized algorithms to
improve performance or runtimes. As a consequence, results can change
between runs, even when provided with the same data and parameters. By
its nature, this randomness renders strict reproducibility impossible.
Debugging is more difficult. If even the small test set (\#9) produces
different results, new users may be able to tell whether the software is
working properly, eroding confidence in the application. When comparing
results between versions or after changing parameters, even small
differences can confuse or muddy the comparison. And especially when
producing results for publications, grants or diagnoses, any analysis
should be absolutely reproducible.

@@

Every package should come with a small test script for users to run
after installation. Its purpose is \emph{not} to check that the software
is working correctly (although that is extremely helpful), but rather to
ensure that it will work at all. This test script can also serve as a
working example of how to run the software, which is valuable in case
its documentation has fallen out of sync with recent changes to the code
itself.

@@

Equally, the test script's output must be easy to interpret. Screens
full of correlation coefficients do not qualify: instead, the script's
output should be something like one line per test, with the test's name
and its pass/fail status, followed by a single summary line saying how
many tests were run and how many passed or failed. If many or all tests
fail because of missing dependencies, that fact should be displayed
once, clearly, rather than once per test, so that users have a clear
idea of what they need to fix and how much work it's likely to take.

@@

Most software has a version number composed of a decimal number that
increments as new versions are released. There are many different ways
to construct and interpret the version, but most importantly for us, a
particular software version run with the same parameters should give
identical results no matter when it's run. Results include correct
output as well as any errors, whether they arise from incorrect input or
are bugs. Every time you release your software, i.e.~distribute it to
someone other than yourself and/or the development team, you should
increment your version number.

\withurl{Semantic versioning}{http://semver.org/} is one of the most common
types of versioning for open-source software. Version numbers take the
form of \emph{MAJOR.MINOR{[}.PATCH{]}}, e.g.~0.2.6-RC1. The major and
minor numbers versions are almost always provided. Changes in the major
version number herald significant changes in the software that are not
backwards compatible, such as changing or removing features or altering
the primary functions of the software. Increasing the minor version
represents incremental improvements in the software like adding new
features. Following the minor version number can be an arbitrary number
of project-specific versions, including patches, builds and qualifiers.
Common qualifiers include \texttt{-SNAPSHOT}, for applications that are
not yet stable or released, and \texttt{-RC} for release candidate prior
to official release.

>>>>

As a relatively young field, bioinformatics is full of newly developed software.
\fixme{MT}{examples here}
Efforts such as
\withurl{the ELIXIR tools and data registry}{http://dx.doi.org/10.1093/nar/gkv1116}
and the
\withurl{Bioinformatics Links Directory}{https://bioinformatics.ca/links\_directory/}
\cite{brazas2012}
have made efforts towards cataloguing it: as of May 2016, the former
has 2500 entries and the latter 1700. Those numbers constantly
grow as both trainees and experienced practitioners produce new
software to support their research.

Typically, this software is developed by one person and 
produces excellent results in their hands. But what happens after the
student leaves someone else from their group wants to use the program? Everyone with
a few years of experience feels a tremor of fear when told, ``Use
Graduated Student's code to analyze your
data''. Often, that software will be undocumented and work in unexpected
ways, if it works at all, relying on nonexistent paths or resources, tuned for
the dataset, or simply an older version than was used in the paper.
The poor student or staff member then has two choices: hack the existing code
to make it work for her, or start over.

@@

\begin{itemize}
\item
  Works for users other than the original author
\item
  Is kept under version control.
\item
  Can be installed on more than one computer or account with relative ease
\item
  Has well-defined input and output formats
\item
  Has documentation that describes what its dependencies are, how to
  install it, and what the options are.
\item
  Comes with enough tests to show that it actually runs.
\end{itemize}

These rules need not be applied to \emph{every} coding effort. The vast majority of
code produced in the marathon of a graduate thesis is used once to answer a
specific question related to a specific dataset. Scientific endeavours often
require speed and rapid iteration over a question, not thoughtful software
engineering and thorough documentation. A README is of no use if you are the
sole user of your code. As the saying goes, not everything worth doing is worth
doing right away. However, once your script is dusted off and run three or four
times for slightly different purposes, when it comes time to publish your
methods, to leave a project, or to pass it on to someone else, it may be time
to apply ten simple rules for robust software.

These simple rules are also necessary steps toward creating a reusable library that
can be shared through a site like CPAN or CRAN. They can apply to both
closed-source and open-source software.  They are generic and can be applied to all
languages, libraries, packages, documentation styles, and
operating systems.  Whether the aim is as simple as sharing the code
with collaborators or as complex as using the software in a production
analysis environment, increasing the robustness of your software
decreases headaches all around.

@@

The README is the first stop for most new users.
At a minimum, it needs to get a new user started and point them towards more
help, if they need it.
A working example using test data (per
\#9) is always appreciated.

\textbf{Explain what the software does:} At the beginning of the README,
explain what the software does in one or two sentences.
There's nothing more frustrating
than spending the time to download and install some software only to
find out that it doesn't do what you thought it did.

\textbf{List required dependencies:} Often, software depends on very
specific versions of libraries, modules, or operating systems. This is
entirely reasonable as long as it is properly documented. Often,
multiple libraries exist with the same or very similar names, so
either provide the commands necessary to download the dependencies or
link to the software homepage. Include the version number for each
dependency.

If your dependency is not publicly available,
you have several options to ensure your software remains useable.
You can add plain-text or code directly to
your repository with appropriate attribution. If the dependency is a binary, use a
\withurl{binary repository manager}{https://en.wikipedia.org/wiki/Binary\_repository\_manager}
such as Artifactory or ProGet. These managers keep versioned copies of
software at constant URLs so they can be downloaded as long as the
manager continues to run. If the dependency must remain closed, place it at an
internal location on shared disk, remove all write permissions, and
link to it from your README, although this method is 
discouraged because of the potential for sharing and risks accidental removal.

\textbf{Installation instructions:} If the software needs to be compiled
or installed, list those instructions in the README. Also
mention if you recommend they use a pre-compiled binary instead
through a system such as pip, yum or apt.

\textbf{Input and output files}: All possible input and output files
should be listed in this section. Do the files conform to a particular
standard, an extension of an existing format, or is it your own
format? If using a standard format, link to the specification and
version. If you extend the standard or have your own format, define it
explicitly, listing all the required fields and acceptable values.
(You get bonus points if you include a script to convert between
a standard format and your file format). If there is no rigorous format
(such as with log files), show an example file, or the first few lines,
and explain what the sections mean.

Input files and their formats are included in most documentation, but
intermediate, auxilliary and log files are often missing. 
\emph{All} files should be listed in the README, even those considered self-explanatory. Log
files are often full of valuable information that can be
mined for the user's specific purpose. Sometimes your users will need to answer a question of
the format, ``Does Software tell you the percentage of reads trimmed to remove
adapter sequences?'' and they can check the README and confidently
say ``yes, it is in the log file''.

\textbf{Attributions and licensing:} Attributions are how you credit
your main contributors; licenses are how you want others to use and
credit your software. By putting it in the README, it is readable \emph{before}
the software is installed (or even downloaded).
Leave no
question in anyone's mind about whether your software can be used
commercially, how much modification is permitted, and how other software
needs to credit you. If your software is not open source, state that clearly.
Attributions can also contain a list of `expert' users
that can be contacted if new users have problems with the software. (for
better or for worse)

\begin{itemize}
\item
  \fixme{GW}{example from real software:\\ https://github.com/dib-lab/khmer/blob/master/README.rst}
\end{itemize}

@@

Usage information provides the first line of help for both first-time and
experienced users of command-line applications.
Ideally, usage is a terse, informative command-line help message that
guides the user in the correct use of your software. Terseness is
important: usage that extends for multiple screens is a nuisance, especially when
printed to standard error instead of standard output (where it can
easily be paged).

Usage should provide all of the information necessary to run the
software. It is invoked either by running the software without
any arguments; running the software with incorrect arguments; or by
explicitly choosing a help or usage option.

@@

\textbf{The syntax for running the program}: This includes the name of the
program and defines the relative
location of optional and required flags, arguments and values for execution.
Arguments in {[}square brackets{]}
are usually optional. An ellipsis (\ldots e.g. ``{[}OPTION{]}\ldots{}'')
indicates that more than one value can be provided.

\textbf{Description}: Similar to the README, the
description reminds users of the software's primary function.

\textbf{Most commonly used arguments, a description of each, and the
default values}: Not all arguments need to appear in the usage, but the most
commonly used ones should be listed here. Users will rely on this for
quick reference when working with your software.

@@

\textbf{Printed to standard output} : So that it can be piped into
\texttt{less}, searched with \texttt{grep}, or compared to the previous
version with \texttt{diff}.

\textbf{Exit with an appropriate exit code}: When usage is invoked by
providing incorrect arguments, the program should exit with a non-zero
code to indicate an error. However, when help is explicitly requested,
the software should not exit with an error.

@@

Root (also known as ``superuser'' or ``admin'') is a special account on
a computer that has (among other things) the power to modify or delete
system files and user accounts. Conversely, files and directories owned
by root usually cannot be modifed by normal users.

@@

Useful parameters are those values that a user will need to modify to suit
their computer, dataset or application. 
The full list of useful parameters is software-specific and so cannot be
detailed here, but here is a short list:

@@

Providing parameters on the
command line increases the flexibility and usability of the program. You
may have determined early in development that 0.58 is an optimal seed for your
original dataset, but that doesn't mean that is the best seed for every
case. Being able to change parameters on the fly to determine if and how
they change the results is important as your software gains more users,
facilitating exploratory analysis and parameter sweeping. Keep in mind
that if a parameter can be adjusted, users will want to be able to turn
off the feature entirely to have a baseline comparison.

@@

You can set reasonable default values
as long as any command line arguments
override those values. Configuration files should be used in preference
to hardcoding the defaults directly. Configuration files can be in a standard
location, e.g. \texttt{.packagerc} in the user's home directory or
provided on the command line as an additional argument.

We caution against overusing configuration files. When a user needs to
locate, open, change and save a file in order to change a parameter, the
import of the change seems larger and discourages experimentation. 
Only values that are unlikely to
change between runs belong in the config file, such as dependencies,
servers, version numbers, network drives, and any other defaults for
your lab or institution. 
Since
all parameters should be echoed in the results, config files should be
cleaned up after execution completes or they occupy valuable disk space.
Use a default configuration file and specify all other parameter values on the
command line.

@@

Hard-coding file paths in a program also makes the software harder to run
in other environments. If your package is installed on a cluster, for
example, the user's data will almost certainly \emph{not} be in the same
directory as the software, and the folder
\texttt{C:\textbackslash{}users\textbackslash{}yourname\textbackslash{}}
will probably not even exist.

@@

In the spirit of code reuse and interoperability, developers often want
to reuse software written by others. 
With a few lines, a call
is made out to the other program, the results are incorporated into the
primary script. Using popular projects reduces the amount of code that
needs to be maintained and adds the strength of vetted software to the
final program.

@@

The way that the second program is invoked can throw up errors.
The program may not be in the user's path, or it may be an
older or newer version and produce results different than expected.
Windows users will be frustrated if you invoke bash or sh explictly or
use shell-specific conventions like `*' expansion.
Even Linux-standard functions available vary slightly between
installs. For example: GNU \texttt{sort} is available on almost every
*nix distribution, but sorts differently depending on locale.

We fully support the reuse of software between projects and have some
suggestions to reduce the aforementioned pains. First, 
make sure that you really need the call-out. If you are
executing GNU sort instead of figuring out how to sort lists in Python,
it may not be worth the pain of integration.

Second, ensure the
appropriate software and version is available. Either allow the user to
configure the exact path to the package, distribute the program with the
dependent software, or download it during installation using a
dependency management system. Regardless, check whether the program is
executable and what version is running. Be sure to remind users in the
documentation what the compatible versions are.

Finally, to ensure support on as many different operating systems as
possible, use native functions for starting other processes, such as
Java's Runtime.exec call, Python's subprocess module, and Perl's system
command, and be sure to capture and report the standard error output of
the subprocess to facilitate debugging.

@@

Many bioinformatics applications rely on randomized algorithms to
improve performance or runtimes. As a consequence, results can change
between runs, even when provided with the same data and parameters. By
its nature, this randomness renders strict reproducibility impossible.
Debugging is more difficult. If even the small test set (\#9) produces
different results, new users may not be able to tell whether the software is
working properly. When comparing
results between versions or after changing parameters, even small
differences can confuse or muddy the comparison. And especially when
producing results for publications, grants or diagnoses, any analysis
should be absolutely reproducible.

@@

Every package should come with a small test script for users to run
after installation. Its purpose is \emph{not} to check that the software
is working correctly (although that is extremely helpful), but rather to
ensure that it will work at all. This test script can also serve as a
working example of how to run the software.

@@

Equally, the test script's output must be easy to interpret. Screens
full of correlation coefficients do not qualify: instead, the script's
output should be simple to understand for non-experts, such as one line per test, with the test's name
and its pass/fail status, followed by a single summary line saying how
many tests were run and how many passed or failed. If many or all tests
fail because of missing dependencies, that fact should be displayed
once, clearly, rather than once per test, so that users have a clear
idea of what they need to fix and how much work it's likely to take.

@@

Most software has a version number composed of a decimal number that
increments as new versions are released. There are many different ways
to construct and interpret the version, but most importantly for us, a
particular software version run with the same parameters should give
identical results no matter when it's run. Results include both correct
output as well as any errors.
Every time you release your software, i.e.~distribute it to
someone other than yourself and/or the development team, you should
increment your version number.

\withurl{Semantic versioning}{http://semver.org/} is one of the most common
types of versioning for open-source software. Version numbers take the
form of \emph{MAJOR.MINOR{[}.PATCH{]}}, e.g.~0.2.6. The major and
minor numbers versions are almost always provided. Changes in the major
version number herald significant changes in the software that are not
backwards compatible, such as changing or removing features or altering
the primary functions of the software. Increasing the minor version
represents incremental improvements in the software, like adding new
features. Following the minor version number can be an arbitrary number
of project-specific identifiers, including patches, builds and qualifiers.
Common qualifiers include \texttt{alpha}, \texttt{beta}, and \texttt{SNAPSHOT}, for applications that are
not yet stable or released, and \texttt{-RC} for release candidates prior
to official release.
